---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(ggplot2)
```


```{r}
n = 1
my_list=list()
MI = "LIGnPt-"
d = 1
for (i in 1:n) {
  for (j in 1:3){
  path = paste0(MI,i,".",j,".xlsx") ###May need to change based on new data file name
  data = read_excel(path)[-1,]
  r = ncol(data)/2
  colnames(data) = paste0(rep(c("V","A"),r),rep(1:r,each = 2))
  my_list[[d]] = as.data.frame(sapply(data, as.numeric))
  d = d+1
  }
}

```

### If required, combine N excel files to 1 excel file with N sheet, which maybe easy to save and share.
```{r}
require(openxlsx)
write.xlsx(my_list, file = "example cv data.xlsx")
```

### The R data file, which is ready to run the selection algorithm 
```{r}
saveRDS(my_list,"my_list.rds")
```

## Find the maximum and minmum current, used for the figure parameters
```{r}
Amax = max(unlist(my_list))
Amin = min(unlist(my_list))
Amax_int = 100*round(Amax/100)
Amin_int = 100*round(Amin/100)
```

## Some data parameters, including the number of electrodes, number of scans, the length of potential vector
```{r}
N = length(my_list) ##Number of electrodes
r = ncol(my_list[[1]])/2 ##Number of scans
d = nrow(my_list[[1]]) ##Length of potential vector (V)
d1 = d/2+1 ##The medium of potential vector
```


## Example plot of last three scans of one electrodes
```{r}
data = my_list[[1]] ##Change the number here to plot for different electrodes
visualize_data = data.frame(A = c(data$A8,data$A9,data$A10),V =rep(data$V1,3),scans = as.character(rep(8:10,each = d)))
p = ggplot(visualize_data)+geom_point(aes(V,A,color = scans),show.legend = T)+ylab("Current (muA)")+xlab("Potential (V)")+scale_y_continuous(breaks=c(Amin_int,Amin_int/2,0,Amax_int/2,Amax_int),limits = c(Amin,Amax))+theme(axis.text.y  = element_text(face="bold"),axis.text.x=element_text(face="bold") )+theme(axis.title.x = element_text(face="bold",size = 14))+theme(axis.title.y = element_text(face="bold",size = 14))
p

```

